discussion

### 2018-05-16 man/ms_3_discussion.Rmd
update

When we explored the chracteristics of the southern sites, we firstly observed a difference in elevation. @Bhuyan2017, in a recent work across the European continent, found a clearly effect of elevation on the resistance to drought of several tree-species, where stands located at higher elevations were less drought sensitive. The higher values of tree growth that we observed in CA-High site could be related with lower water stress at high-elevation sites, since moisture availability is a key factor limiting tree growth for *Q. pyrenaica* [@GeaIzquierdo2014]. In fact, the differences in available water of soils could explain our results. [@CoboDiaz2017], who studied microbiota of this location in an elevational gradient, found higher values of available water of the forest soils located on high elevation compared with those located at low elevations. In this sense, a remarkably observation supporting this idea, is the presence of a traditional irrigation ditches located uphill from the CA-High site. In most of the high-mountain watersheds of Sierra Nevada, there is a system of historical irrigation channels, know as *acequias de careo*, that was used since Middle Age to cultivated these valleys [@MartinCivantos2014; @MartinCivantos2016]. These ditches run through the hillsides of valleys releasing water through several points that filtrate the water to recharge the aquifer and then irrigates the foot of the slopes [@MartinCivantos2014; @Jodar2017]. There is a ditch (*Acequia de la Era Alta*) located uphill the CA-High site, which functions from March to June [@RuizRuiz2017]. It brings water from snowmelt and from a nearby stream, soaking the *Q. pyrenaica* forests. 
Finally another idea to consider is the different degree of anthropic impact related to elevation.

>>>>

When we explored the chracteristics of the southern sites, we firstly observed a difference in elevation. @Bhuyan2017, in a recent work across the European continent, found a clearly effect of elevation on the resistance to drought of several tree-species, where stands located at higher elevations were less drought sensitive. The higher values of tree growth that we observed in CA-High site could be related with lower water stress at high-elevation sites, since moisture availability is a key factor limiting tree growth for *Q. pyrenaica* [@GeaIzquierdo2014]. In fact, the differences in available water of soils could explain our results. @CoboDiaz2017, in a work analyzing the variation of the microbiota along an elevational gradient within this location, found higher values of available water of the forest soils located on high elevation compared with those located at low elevations. In this sense, a remarkable observation is the presence of traditional irrigation ditches which could provide an additional water supply. In most of the high-mountain watersheds of Sierra Nevada, there is a system of historical irrigation channels, know as *acequias de careo*, that was used since Middle Age to cultivated these valleys [@MartinCivantos2014; @MartinCivantos2016]. These ditches run through the hillsides of valleys releasing water through several points that filtrate the water to recharge the aquifer and then irrigates the foot of the slopes [@MartinCivantos2014; @Jodar2017]. The network of traditional ditches present in one of our sampling site (Cáñar) was recentlty described, with detailed information about water use [@RuizRuiz2017] and its hydrological functioning [@MartinMontanes2015]. There is a ditch (*Acequia de la Era Alta*) located uphill the CA-High site, which functions from March to June [@RuizRuiz2017]. It brings water from snowmelt and from a nearby stream, soaking the *Q. pyrenaica* forests. This traditional system could supply an extra of water that could be used by the trees located downstream of these channels. This extra of water is particularly important for trees of this dry-rear edge and could explain the higher values of BAI for trees located in CA-High site.

Finally another idea to consider is the different degree of anthropic impact related to elevation.
